---
title: "Bayes factor analysis example"
author: "Holly Zaharchuk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highligh: github
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# knitr settings
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

I calculated Bayes factors (BFs) corresponding to the four traditional ANOVAs in each time window (EAN, P600) and distribution (midline [Fz, Cz, Pz], lateral [LF, RF, LP, RP]). Below, I show the calculation for the EAN effect along the midline. 

[This article](https://psycnet.apa.org/doiLanding?doi=10.1037/met0000057) was incredibly helpful for understanding how to implement the analysis with the `BayesFactor` package, and [this open access article](https://link.springer.com/article/10.3758/s13423-017-1323-7) provides useful information on interpretation (see Table 1). 

```{r, eval=FALSE}
# Load package for calculating Bayes factors
library(BayesFactor)
```

The relevant variable names are:

- ID = participant ID number
- chlabel = individual electrodes or channels (e.g., Pz)
- group = participant group (Unmarked or Southern)
- cond = sentence condition (DM or SM)
- window = time window (EAN or P600)
- distribution = group of electrodes or channels (midline, lateral)

```{r, eval=FALSE}
# Convert variables to factors (required by anovaBF function)
erp_data <- erp_data %>%
  mutate(ID = as.factor(ID),
         chlabel = as.factor(chlabel),
         group = as.factor(group),
         cond = as.factor(cond))
```

```{r, eval=FALSE}
# Run Bayesian ANOVA
# EAN, midline, interaction among electrodes, groups, and conditions
# anovaBF function also requires random factor for participants
bf_ean_center <- anovaBF(value ~ group*chlabel*cond + ID, 
              data = erp_data %>% 
                dplyr::filter(window == "ean" & 
                              distribution == "center"),
              whichModels="withmain",
              whichRandom = "ID", 
              iterations = 100000)

# Extract all BFs, calculated for every term/interaction
bf_ean_center <- extractBF(bf_ean_center) %>%
  name_rows()
```

Running a Bayesian ANOVA compares every combination of variables (main effects and interactions) to a null model. As a result, the BFs in this analysis don't tell you anything about the probability of one model relative to another. 

In order to calculate a BF comparing two models, you need to extract the BFs for the two models you want to compare, and then calculate a ratio. That's it!

```{r, eval=FALSE}
# Pull BF of model for tranditional ANOVA (not including group)
# Preferred model
bf_num <- bf_ean_center %>%
  dplyr::filter(.rownames == "chlabel + cond + chlabel:cond + ID") %>%
  pull(bf)
```

```{r, eval=FALSE}
# Pull BF of model including group
# Alternative model
bf_den <- bf_ean_center %>%
  dplyr::filter(.rownames == "chlabel + cond + chlabel:cond + 
                              group + chlabel:group + cond:group + c
                              hlabel:cond:group + ID") %>%
  pull(bf)
```

The ratio between the BFs of the preferred and alternative models provides quantitative evidence for/against the model in the numerator relative to the model in the denominator

For example, if the ratio equals 2, then the preferred model has a 2-to-1 likelihood of producing the observed data relative to the alternative model.

In other words, while traditional hypothesis testing yields a binary significance test for a model, Bayesian hypothesis testing yields a continuous measure of support for/against a model.

```{r, eval=FALSE}
# Calculate BF for comparison between models with/without group factor
bf_ean_center_comp <- bf_num / bf_den
```

This ratio (preferred/alternative) would be reported as $BF_{10}$, while the opposite ratio (alternative/preferred) would be reported as $BF_{01}$.